An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-04-30 EconGrowth-R-Functions.Rnw
Edited the text for the CD and CDe parameter estimation section.

#################
# Creates a plot with y residuals vs. energy (e). The y residuals are from the specified modelType without energy.
# modelType: one of "CD" or "CES"
# energyType: one of "Q", "X", or "U" to serve as the ordinate of the graph
# countryAbbrev: set to NA (the default) if you want a lattice graph with all countries shown. Set to one of
#                US, UK, JP, CN, ZA, SA, IR, TZ, ZM for an individual graph.
## 
createPartialResidualPlot <- function(modelType, energyType, countryAbbrev=NA, textScaling=1.0){
  xLabelString <- paste("$", tolower(energyType), "$", sep="")
  yLabelString <- paste("$y$ residuals (", modelType, " without energy)", sep="")
  if (! is.na(countryAbbrev)){
    # An individual graph for a single country is desired.
    data <- calcResiduals(modelType=modelType, countryAbbrev=countryAbbrev, energyType=energyType)
    plot <- xyplot(resid ~ iEToFit, data=data,
                   type=c("p"),
                   col="black",
                   scales=list(cex=scaleTextSize*textScaling, #controls text size on scales.
                               tck=scaleTickSize), #controls tick mark length. < 0 for inside the graph.
                   xlab=list(label=xLabelString, cex=textScaling),
                   ylab=list(label=yLabelString, cex=textScaling)
                   )
  } else {
    # Lattice graph will all countries is desired.
    data <- do.call("rbind", lapply(countryAbbrevsAlph, calcResiduals, modelType=modelType, energyType=energyType))
    # If useful work is desired, need to trim the data set.
    if (energyType == "U"){
      # Use data from only those countries that have U defined.
      data <- subset(data, !is.na(iEToFit))
      layout <- threePanelLayoutSpec
      factorLevels <- countryNamesAlphU # Want only those countries for which U is known
      indexCond <- list(countryOrderForGraphsU) # Orders as US, UK, JP
    } else {
      # Use data from all countries, because all countries have Q and X data.
      factorLevels <- countryNamesAlph # We want all countries shown
      layout <- ninePanelLayoutSpec
      indexCond <- list(countryOrderForGraphs) # Orders as US, UK, JP, CN, ZA, SA, IR, TZ, ZM
    }
    plot <- xyplot(resid ~ iEToFit | Country, data=data,
                   type = c("p"),
                   index.cond = indexCond, #orders the panels.
                   layout = layout,
                   strip = strip.custom(factor.levels=factorLevels, 
                                        bg="transparent", # Sets background transparent to match the graph itself.
                                        par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                        ),
                   as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                   col = "black",
                   scales = list(cex=scaleTextSize*textScaling, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 alternating=FALSE), # prevents left/right, top/bot alternating
                   xlab = list(label=xLabelString, cex=textScaling),
                   ylab = list(label=yLabelString, cex=textScaling)
                   )
  }
  return(plot)
}

>>>>

#################
# Creates a plot with y residuals vs. energy (e). The y residuals are from the specified modelType without energy.
# modelType: one of "CD" or "CES"
# energyType: one of "Q", "X", or "U" to serve as the ordinate of the graph
# countryAbbrev: set to NA (the default) if you want a lattice graph with all countries shown. Set to one of
#                US, UK, JP, CN, ZA, SA, IR, TZ, ZM for an individual graph.
## 
createPartialResidualPlot <- function(modelType, energyType, countryAbbrev=NA, textScaling=1.0){
  xLabelString <- paste("$", tolower(energyType), "$", sep="")
  yLabelString <- paste("$y$ residuals (", modelType, " without energy)", sep="")
  if (! is.na(countryAbbrev)){
    # An individual graph for a single country is desired.
    data <- calcResiduals(modelType=modelType, countryAbbrev=countryAbbrev, energyType=energyType)
    plot <- xyplot(resid ~ iEToFit, data=data,
                   type=c("p"),
                   col="black",
                   scales=list(cex=scaleTextSize*textScaling, #controls text size on scales.
                               tck=scaleTickSize), #controls tick mark length. < 0 for inside the graph.
                   xlab=list(label=xLabelString, cex=textScaling),
                   ylab=list(label=yLabelString, cex=textScaling)
                   )
  } else {
    # Lattice graph will all countries is desired.
    data <- do.call("rbind", lapply(countryAbbrevsAlph, calcResiduals, modelType=modelType, energyType=energyType))
    # If useful work is desired, need to trim the data set.
    if (energyType == "U"){
      # Use data from only those countries that have U defined.
      data <- subset(data, !is.na(iEToFit))
      layout <- threePanelLayoutSpec
      factorLevels <- countryNamesAlphU # Want only those countries for which U is known
      indexCond <- list(countryOrderForGraphsU) # Orders as US, UK, JP
    } else {
      # Use data from all countries, because all countries have Q and X data.
      factorLevels <- countryNamesAlph # We want all countries shown
      layout <- ninePanelLayoutSpec
      indexCond <- list(countryOrderForGraphs) # Orders as US, UK, JP, CN, ZA, SA, IR, TZ, ZM
    }
    xLimits <- c(0.5, 4.5) # Maintain consistent x-axis limits across all graphs
    yLimits <- c(-0.2, 0.2) # Maintain consistent y-axis limits across all graphs
    plot <- xyplot(resid ~ iEToFit | Country, data=data,
                   type = c("p"),
                   index.cond = indexCond, #orders the panels.
                   layout = layout,
                   strip = strip.custom(factor.levels=factorLevels, 
                                        bg="transparent", # Sets background transparent to match the graph itself.
                                        par.strip.text=list(cex=textScaling) # Scales text in the strip.
                                        ),
                   as.table = TRUE, #indexing of panels starts in upper left and goes across rows.
                   col = "black",
                   xlim=xLimits, 
                   ylim=yLimits, 
                   scales = list(cex=scaleTextSize*textScaling, #controls text size on scales.
                                 tck=scaleTickSize, #controls tick mark length. < 0 for inside the graph.
                                 alternating=FALSE), # prevents left/right, top/bot alternating
                   xlab = list(label=xLabelString, cex=textScaling),
                   ylab = list(label=yLabelString, cex=textScaling)
                   )
  }
  return(plot)
}

### 2013-04-30 EconGrowth.Rnw
Edited the text for the CD and CDe parameter estimation section.

\item{Heun: Adjust variable threePanelGraphHeight in the Functions file to produce U partial residual plots that are exactly the same size as the graphs in the 9-panel plots.}

@@

@@

\noindent from which $\beta = 1 - \alpha$ can be found. The above reparameterization ensures the $\alpha + \beta = 1$ for constant returns to scale. ****** Say something here about constraints when we get them figured out. *********** $\lambda$ and $\alpha$ were estimated for all countries with the \citet{R} function \texttt{nls}. Figure \ref{fig:CD_Params_Graph} shows values of and 95\% confidence intervals for $\lambda$ and $\alpha$. Appendix \ref{sec:Model_Parameters} contains tables with numerical values of model parameters and confidence intervals.

@@

To estimate the parameters $\lambda$, $\alpha$, $\beta$, and $\gamma$ in the energy-augmented Cobb-Douglas model, we reparameterized Equation \ref{eq:CDe} as

@@

\noindent from which $\alpha = \min(a,b)$, $\beta = |b-a|$, and $\gamma = \max(a,b)$ can be calculated. By constraining $a \in [0,1]$ and $b \in [0,1]$, we ensure that $\alpha \in [0,1]$, $\beta \in [0,1]$, and $\gamma \in [0,1]$. The \citet{R} function \texttt{nls} was used to estimate values for $\lambda$, $a$, and $b$.

>>>>

@@

\item{\st{Heun: Ensure consistent x and y axis limits in partial residual plots}}
\item{\st{Heun: Use full country names instead of 2-letter abbreviations in partial residual plots}}
\item{\st{Heun: Adjust variable threePanelGraphHeight in the EconGrowth-R-Functions.Rnw file to produce U partial residual plots that are exactly the same size as the graphs in the 9-panel plots.}}

@@

\noindent to ensure $\alpha + \beta = 1$ for constant returns to scale. $\lambda$ and $\alpha$ were estimated for all countries in Equation \ref{eq:CD_reparam} by the \citet{R} function \texttt{nls}. By constraining $\alpha \in [0,1]$ during the fitting process, we ensure that $\beta \in [0,1]$ also. ****** Is this truly how we implemented the constraints? Finish after we actually implement the constraints. *********** The value of $\beta$ can be found with $\beta = 1 - \alpha$. Figure \ref{fig:CD_Params_Graph} shows values of and 95\% confidence intervals for $\lambda$ and $\alpha$. Appendix \ref{sec:Model_Parameters} contains tables with numerical values of model parameters and confidence intervals.

@@

To estimate the parameters $\lambda$, $\alpha$, $\beta$, and $\gamma$ in the energy-augmented Cobb-Douglas model, Equation \ref{eq:CDe} was reparameterized as

@@

\noindent to ensure that $\alpha + \beta + \gamma = 1$. By constraining $a \in [0,1]$ and $b \in [0,1]$, we ensure that $\alpha \in [0,1]$, $\beta \in [0,1]$, and $\gamma \in [0,1]$. Given values of $a$ and $b$ that represent the least-squares parameter optimization for a given economy, $\alpha = \min(a,b)$, $\beta = |b-a|$, and $\gamma = \max(a,b)$ can be calculated. The \citet{R} function \texttt{nls} was used to estimate the least-squares values of $\lambda$, $a$, and $b$.
